---
title: "XXXXXXXX"
author: "Nikolay Petrov"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
---

# SETUP

The code here is an extension of the Python codebook (`codebook.ipynb`) so to see how the data that used here has come to be, see that notebook.

```{r include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9)
```

```{r}
show_data_markdown <- function(data, group_var=NULL, rows_by_group=Inf, row_limit=Inf) {
  if (!is.null(group_var)) {
    data_to_display <- data %>%
      group_by(.data[[group_var]])
  } else {
    data_to_display <- data
  }
  data_to_display <- data_to_display %>%
    filter(row_number()<=rows_by_group) %>%
    ungroup() %>%
    filter(row_number()<=row_limit)
  
  if (isTRUE(getOption('knitr.in.progress'))) {
    return(knitr::kable(data_to_display))
  } else {
    return(data_to_display)
  }
}
```

Packages loading and setup:

```{r warning=FALSE, message=FALSE}
options(dplyr.summarise.inform = FALSE)

packages <- c("ggplot2",
              "ggpubr",
              "tidyverse",
              "lavaan",
              "semPlot",
              "psych",
              "GPArotation",
              "EFA.dimensions",
              "lavaan",
              "semPlot",
              "semTools",
              "data.table"
              )

if (!all(packages %in% (.packages()))) {
  #using invisible to hide output
  invisible(lapply(packages,
         FUN = function(x) {
           if (!require(x, character.only = TRUE)) {
             install.packages(x, dependencies = TRUE)
           }
           library(x, character.only = TRUE)
         }))
}

theme_set(theme_classic())
theme_update(plot.title = element_text(size=25, hjust=0.5),
             plot.subtitle = element_text(size=18, hjust=0.5, face="italic"),
            axis.title = element_text(size=15),
            axis.text = element_text(size=10),
            legend.title = element_text(size=20),
            legend.text = element_text(size=15),
            legend.position = "bottom",
            legend.justification="right")
```

```{r}
# Custom formatting function for APA style
format_apa <- function(x, is_p_val=F) {
  rounding_digits <- 2
  
  if (is_p_val) {
    if (x<0.001) {
      return ("< .001")
    }
    rounding_digits <- 3
  }
  
  sign <- ifelse(x < 0, "-", "")
  formatted <- formatC(abs(x), format = "f", digits = rounding_digits)
  formatted <- gsub("^0", "", formatted) # Remove leading zero
  formatted <- paste0(sign, formatted)
  
  if (is_p_val) {
    return(paste0("= ", formatted))
  }
  return(formatted)
}

# Custom label mappings
model_labels <- c("serapiogarcia" = "Serapio-García et al. (2023)", "bbc" = "BBC Human data",
                  "generic_gpt35" = "Generic personas\nGPT3.5", "generic_gpt4" = "Generic personas\nGPT4", 
                  "silicon_gpt35" = "Silicon personas\nGPT3.5", "silicon_gpt4" = "Silicon personas\nGPT4")
model_labels2 <- c("gpt35" = "GPT3.5", "gpt4" = "GPT4")

scale_labels <- c("BFI" = "BFI", "BPAQ" = "BPAQ", "SSCS" = "SSCS", "PANAS" = "PANAS")
```

# Figure: internal consistency

```{r fig.width=16}
data <- read_csv(file.path("data", "data_for_R", "combined_cronbach_alpha_df.csv")) %>%
  pivot_longer(cols=-c("scale", "dimension"),
               names_to="model",
               values_to="alpha") %>%
  mutate(model = gsub("_alpha", "", model),
         scale = recode(scale, !!!scale_labels),
         model = recode(model, !!!model_labels),
         model = factor(model),
         # model = fct_relevel(model, levels(model)[3], after=1)
         )

p_bfi <- data %>%
  filter(scale=="BFI") %>%
  ggplot(aes(x = dimension, y = alpha)) +
    geom_col(position = position_dodge()) +
    geom_hline(yintercept=0.7) +
    geom_text(aes(label = format_apa(alpha)), 
              position = position_dodge(width = 0.9), 
              vjust = -0.3, 
              color = "black", 
              size = 3.5, 
              check_overlap = TRUE) +
    facet_wrap(~scale+model, scales = "free_x", nrow=4, ncol=1) +
    scale_y_continuous(limits=c(0, 1.03)) +
    theme(strip.text = element_blank(),
          axis.text.x = element_text(angle=45, hjust=1))

p_others <- data %>%
  filter(scale!="BFI") %>%
  ggplot(aes(x = dimension, y = alpha)) +
    geom_col(position = position_dodge()) +
    geom_hline(yintercept=0.7) +
    geom_text(aes(label = format_apa(alpha)), 
              position = position_dodge(width = 0.9), 
              vjust = -0.3, 
              color = "black", 
              size = 3.5, 
              check_overlap = TRUE) +
    facet_wrap(~model+scale, scales = "free_x", nrow=4) +
    scale_y_continuous(limits=c(0, 1.03)) +
    theme(strip.text = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          axis.line.y = element_blank())

ggarrange(p_bfi, p_others,
          labels = c("", ""),
          # nrow=1, ncol=2,
          widths=c(0.25, 0.75)
          # common.legend=T,
          # legend="bottom"
          )
```

```{r fig.width=24}
p <- data %>%
  ggplot(aes(x = dimension, y = alpha)) +
    geom_col(position = position_dodge()) +
    geom_hline(yintercept=0.7) +
    geom_text(aes(label = format_apa(alpha)), 
              position = position_dodge(width = 0.9), 
              vjust = -0.3, 
              color = "black", 
              size = 3.5, 
              check_overlap = TRUE) +
    # facet_wrap(~scale+model, scales = "free_x") +
    scale_y_continuous(limits=c(0, 1.03))

p + facet_grid(rows = vars(model), cols = vars(scale), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0)
  ) +
  labs(x="Dimension", y="Cronbach's alpha")
```

# Figure: Criterion Validity

```{r fig.width=12}
data <- read_csv(file.path("data", "data_for_R", "combined_validity_df.csv")) %>%
  pivot_longer(cols=-c("var1", "var2"),
               names_to="model",
               values_to="r") %>%
  rowwise() %>%
  mutate(var1 = if_else(startsWith(var1, "BFI_"), strsplit(var1, "_")[[1]][2], var1)) %>%
  ungroup() %>%
  mutate(model = gsub("_r", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model),
         model = fct_relevel(model, levels(model)[3], after=0)
         )

data
```

```{r fig.width=20}
p <- ggplot(data, aes(x = var2, y = r)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(r)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) #+
    # facet_wrap(~var1, scales = "free_x")

p + facet_grid(rows = vars(model), cols = vars(var1), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0),
    axis.title.x = element_blank()
  ) + 
  labs(y="Correlation (Pearson's r)")
```

# Figure: Intercorrelations

```{r}
# Read the data
data <- read.csv(file.path("data", "data_for_R", "intercorrs.csv")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         # model = fct_relevel(model, levels(model)[3], after=1)
         )

data
```

```{r fig.width=9, fig.height=12}
ggplot(data, aes(x = var1, y = r, fill=model)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(r)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) +
    facet_wrap(~var2, ncol=1)
```

```{r fig.width=30}
p <- ggplot(data, aes(x = var1, y = r)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(r)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) #+
    # facet_wrap(~var2, ncol=1)

p + facet_grid(rows = vars(model), cols = vars(var2), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0),
    axis.title.x = element_blank()
  ) + 
  labs(y="Correlation (Pearson's r)")
```

# Figure: Item-level frequencies

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_bfi_item_frequencies.csv")) %>%
  mutate(model = gsub("_relative_frequency", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model),
         # model = fct_relevel(model, levels(model)[3], after=1)
         )
data
```

```{r fig.width=12}
p <- ggplot(data, aes(x = response_reversed, y = relative_frequency)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(relative_frequency)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE)

p + facet_grid(rows = vars(model), cols = vars(dimension), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0)
  ) + 
  labs(y="Frequency", x="Item-level response") +
  scale_y_continuous(limits=c(0, 0.8))
```

```{r fig.width=12, fig.height=9}
ggplot(data, aes(x = response_reversed, y = relative_frequency)) +
    geom_col(position = position_dodge()) +
    # geom_text(aes(label = format_apa(r)), 
    #             position = position_dodge(width = 0.9), 
    #             vjust = -0.3, 
    #             color = "black", 
    #             size = 3.5, 
    #             check_overlap = TRUE) +
    facet_wrap(~dimension+model)
```

# EFA

```{r}
process_data <- function(filepath) {
    # Read the CSV file from the given file path
    data <- read_csv(filepath)

    # Apply transformations
    data <- data %>%
        mutate(uid = as.factor(uid),
               dimension = substr(dimension, 1, 1)) %>%
        rename("response" = "response_reversed") %>%
        arrange(dimension, item_index) %>%
        pivot_wider(id_cols = uid,
                    names_from = c("dimension", "item_index"),
                    names_sep = "",
                    values_from = "response")

    # Return the transformed data
    return(data)
}

generic_gpt35_df <- process_data(file.path("data", "data_for_R", "generic_gpt35_df.csv"))
generic_gpt4_df <- process_data(file.path("data", "data_for_R", "generic_gpt4_df.csv"))

silicon_gpt35_df <- process_data(file.path("data", "data_for_R", "silicon_gpt35_df.csv"))
silicon_gpt4_df <- process_data(file.path("data", "data_for_R", "silicon_gpt4_df.csv"))
```

```{r}
get_fa_model_loadings <- function(data, model_label) {
  cor_matrix <- cor(select(data, -uid),
                  use = "pairwise.complete.obs")

  model <- fa(r = cor_matrix, nfactors = 5, 
                       rotate = "oblimin", fm = "pa", 
                       max.iter = 5000, n.obs = length(rownames(data)))
  
  df <- as.data.frame(unclass(model$loadings))
  df$item <- rownames(df)
  rownames(df) <- NULL
  df <- df %>%
    pivot_longer(cols=-item,
                 names_to="factor",
                 values_to="loading") %>%
    mutate(model=model_label)
  
  return(df)
}
```

```{r}
if (!dir.exists(file.path("data", "bbc_data"))) {
  load(file.path("data", "bbc_data_for_sharing", "bbc_loadings.Rdata"))
} else {
  bbc_df <- process_data(file.path("data", "bbc_data", "bbc_df.csv"))
  bbc_loadings <- get_fa_model_loadings(bbc_df, "bbc")
  save(bbc_loadings, file=file.path("data", "bbc_data_for_sharing", "bbc_loadings.Rdata"))
}
```

```{r}
efa_loadings <- bind_rows(bbc_loadings,
                          get_fa_model_loadings(generic_gpt35_df, "generic_gpt35"),
                          get_fa_model_loadings(generic_gpt4_df, "generic_gpt4"),
                          get_fa_model_loadings(silicon_gpt35_df, "silicon_gpt35"),
                          get_fa_model_loadings(silicon_gpt4_df, "silicon_gpt4")
                          ) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         # model = fct_relevel(model, levels(model)[3], after=1)
         )
efa_loadings
```

## Figure EFA loadings

```{r fig.width=15, fig.height=6}
efa_loadings %>%
  mutate(loading = abs(loading)) %>%
  ggplot(aes(x=factor, y=item, fill=loading)) +
      geom_tile() +
      scale_fill_gradient2(low="white", mid="darkgrey", high="black", midpoint=0.5, limit=c(0,1), space="Lab", name="Factor\nLoading") +
      facet_wrap(~model, nrow=1) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(x = "Principal Component", y = "Variable")
```

## Table EFA loadings

```{r}
efa_loadings %>%
  mutate(dimension = str_sub(item, 1, 1),
         item_index = as.double(str_sub(item, 2, -1))) %>%
  arrange(dimension, item_index, model, factor) %>%
  mutate(item=paste0(dimension, item_index)) %>%
  pivot_wider(id_cols=item,
              names_from=c("model", "factor"),
              values_from="loading") %>%
  mutate(across(is.numeric, ~format_apa(.))) #%>%
  # write.csv("figures-tables/efa_loadings.csv", row.names=F)
```

# Figure: Trait bias

```{r}
data <- read_csv(file.path("data", "data_for_R", "bias_df.csv")) %>%
  mutate(model = recode(model, !!!model_labels2))
```

```{r fig.width=12, fig.height=6}
data_summary_stats_labels <- data %>%
  group_by(model, trait) %>%
  summarise(label = paste0('<i>M</i>=', format_apa(mean(bias)), ' (<i>SD</i>=', format_apa(sd(bias)), ')'))

data_p_val_labels <- data %>%
  group_by(trait) %>%
  summarise(label = if_else(
    t.test(bias~model, data=cur_data())$p.value < 0.05,
    paste0('<b><i>p</i> ', format_apa(t.test(bias~model, data=cur_data())$p.value, T), '</b>'),
    paste0('<i>p</i> ', format_apa(t.test(bias~model, data=cur_data())$p.value, T))
    )
  )


ggplot(data, aes(x=model, y=bias)) +
    geom_violin(width=0.6) +
    geom_boxplot(width=0.05, alpha=0.2) +
    ggtext::geom_richtext(data=data_summary_stats_labels, 
                            mapping=aes(y=3.5, x=model, label=label, group=NULL, fill=NULL), 
                            size=3,
                            show.legend=F) +
    ggtext::geom_richtext(data=data_p_val_labels, 
                            mapping=aes(y=3, x=1.5, label=label, group=NULL, fill=NULL), 
                            size=3,
                            show.legend=F) +
    labs(x="Model", y="Bias") +
    facet_wrap(~trait) +
    scale_y_continuous(limits=c(0, 3.7))
```

# Figure: Bias correlations

```{r}
data <- read_csv(file.path("data", "data_for_R", "bias_corrs_df.csv"))
data
```

```{r}
labels_mapper_p1 <- c(
  "BFI_Agreeableness" = "Agreeableness", "BFI_Conscientiousness" = "Conscientiousness",
  "BFI_Extraversion" = "Extraversion", "BFI_Openness" = "Openness", "BFI_Neuroticism" = "Neuroticism"
)
labels_mapper_p2 <- c(
  "age" = "Age", "sex" = "Sex", "ethnic" = "Ethnicity", "income" = "Income",
  "rstat_1" = "Relationship status", "n_sib" = "Number of siblings"
)
labels_mapper_p3 <- c(
  "st_pub" = "State or private schooling", "chldrn" = "Number of children", 
  "occ_sta" = "Occupational status", "occ_cat" = "Occupation category",
  "m_schl" = "Mother's education", "f_schl" = "Father's education"
)
```


```{r fig.width=12, fig.height=9}
p1 <- data %>%
  filter(str_detect(var, "^BFI_")) %>%
  mutate(var = recode(var, !!!labels_mapper_p1)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="Correlation coefficient", fill="") +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1
```

```{r fig.width=12, fig.height=9}
p2_vars <- c("age", "sex", "ethnic", "income", "rstat_1", "n_sib")

labels_mapper <- c("st_pub" = "State or private schooling", "chldrn" = "Number of children", 
                   "occ_sta" = "Occupational status", "occ_cat" = "Occupation category",
                   "m_schl" = "Mother's education", "f_schl" = "Father's education")

p2 <- data %>%
  filter(var %in% p2_vars) %>%
  mutate(var = recode(var, !!!labels_mapper_p2)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="", fill="") +
      guides(fill="none") +
      scale_y_continuous(limits=c(-0.05, 0.11)) +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3 <- data %>%
  filter(!str_detect(var, "^BFI_"), !var %in% p2_vars) %>%
  mutate(var = recode(var, !!!labels_mapper_p3)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="", fill="") +
      # guides(fill="none") +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      scale_y_continuous(limits=c(-0.05, 0.11)) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3
```

```{r fig.width=12, fig.height=9}
p2_p3 <- ggarrange(p2, p3, nrow=2, ncol=1, common.legend = T, legend="none")
p2_p3
```


```{r fig.width=12, fig.height=9}
ggarrange(p1, p2_p3,
          # labels = c("A: Study 1", "B: Study 2", "C: Study 5"),
          font.label=list(size=20, color="black", face="bold"),
          nrow=1, ncol=2,
          common.legend=T,
          legend="bottom")
```

```{r fig.width=12, fig.height=9}
ggplot(data, aes(x = var, y = r, fill=model)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(r)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Figure: Supplementary - Analysis of the first token of the responses

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_first_token_digit_item_frequencies.csv")) %>%
  mutate(model = gsub("_relative_frequency", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model),
         first_token_is_digit = if_else(first_token_is_digit, "first token is digit", "first token is NOT digit")
         )
data
```

```{r fig.width=12}
p <- ggplot(data, aes(x = response_reversed, y = relative_frequency)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(relative_frequency)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE)

p + facet_grid(rows = vars(model), cols = vars(first_token_is_digit), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0)
  ) + 
  labs(y="Frequency", x="Item-level response") +
  scale_y_continuous(limits=c(0, 0.8))
```
























