---
title: "XXXXXXXX"
author: "Nikolay Petrov"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
---

# SETUP

The code here is an extension of the Python codebook (`codebook.ipynb`) so to see how the data that used here has come to be, see that notebook.

```{r include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9)
```

```{r}
show_data_markdown <- function(data, group_var=NULL, rows_by_group=Inf, row_limit=Inf) {
  if (!is.null(group_var)) {
    data_to_display <- data %>%
      group_by(.data[[group_var]])
  } else {
    data_to_display <- data
  }
  data_to_display <- data_to_display %>%
    filter(row_number()<=rows_by_group) %>%
    ungroup() %>%
    filter(row_number()<=row_limit)
  
  if (isTRUE(getOption('knitr.in.progress'))) {
    return(knitr::kable(data_to_display))
  } else {
    return(data_to_display)
  }
}

save_versioned_plot <- function(plot, filename, directory = "figures-tables/", width=12, height=9, ...) {
  # Ensure the directory ends with a slash
  if (!grepl("/$", directory)) {
    directory <- paste0(directory, "/")
  }
  
  # Construct the pattern to search for existing files
  pattern <- paste0("^", filename, "_v[0-9]+\\.png$")
  
  # List all files in the directory
  files <- list.files(directory, pattern = pattern)
  
  # Extract version numbers
  versions <- gsub(pattern = paste0("^", filename, "_v([0-9]+)\\.png$"), replacement = "\\1", x = files)
  versions <- as.integer(versions)
  
  # Determine the next version number
  if (length(versions) == 0) {
    next_version <- 1
  } else {
    next_version <- max(versions) + 1
  }
  
  # Construct the new filename with the next version number
  new_filename <- sprintf("%s%s_v%02d.png", directory, filename, next_version)
  
  # Save the plot with additional arguments passed to ggsave
  ggsave(new_filename, plot = plot, width = width, height = height, ...)
  
  cat(sprintf("Plot saved as %s\n", new_filename))
}
```

Packages loading and setup:

```{r warning=FALSE, message=FALSE}
options(dplyr.summarise.inform = FALSE)

packages <- c("ggplot2",
              "ggpubr",
              "tidyverse",
              "lavaan",
              "semPlot",
              "psych",
              "GPArotation",
              "EFA.dimensions",
              "lavaan",
              "semPlot",
              "semTools",
              "data.table"
              )

if (!all(packages %in% (.packages()))) {
  #using invisible to hide output
  invisible(lapply(packages,
         FUN = function(x) {
           if (!require(x, character.only = TRUE)) {
             install.packages(x, dependencies = TRUE)
           }
           library(x, character.only = TRUE)
         }))
}

theme_set(theme_classic())
theme_update(plot.title = element_text(size=25, hjust=0.5),
             plot.subtitle = element_text(size=18, hjust=0.5, face="italic"),
            axis.title = element_text(size=15),
            axis.text = element_text(size=10),
            legend.title = element_text(size=20),
            legend.text = element_text(size=15),
            legend.position = "bottom",
            legend.justification="right")
```

```{r}
# Custom formatting function for APA style
format_apa <- function(x, is_p_val=F) {
  rounding_digits <- 2
  
  if (is_p_val) {
    if (x<0.001) {
      return ("< .001")
    }
    rounding_digits <- 3
  }
  
  sign <- ifelse(x < 0, "-", "")
  formatted <- formatC(abs(x), format = "f", digits = rounding_digits)
  formatted <- gsub("^0", "", formatted) # Remove leading zero
  formatted <- paste0(sign, formatted)
  
  if (is_p_val) {
    return(paste0("= ", formatted))
  }
  return(formatted)
}

# Custom label mappings
model_labels <- c("serapiogarcia" = "Serapio-García et al. (2023)", "bbc" = "BBC Human data",
                  "generic_gpt35" = "Generic personas\nGPT3.5", "generic_gpt4" = "Generic personas\nGPT4", 
                  "silicon_gpt35" = "Silicon personas\nGPT3.5", "silicon_gpt4" = "Silicon personas\nGPT4")
model_labels2 <- c("gpt35" = "GPT3.5", "gpt4" = "GPT4")

scale_labels <- c("BFI" = "BFI", "BPAQ" = "BPAQ", "SSCS" = "SSCS", "PANAS" = "PANAS")
```

## Reading in LLM data

```{r}
process_data <- function(filepath) {
    # Read the CSV file from the given file path
    data <- read_csv(filepath, show_col_types=F)
    
    # add the scale column to the bbc data for consistency
    if (grepl("bbc_df.csv", filepath, fixed=T)) {
      data <- data %>%
        mutate(scale="BFI")
    }

    # Apply transformations
    data <- data %>%
        mutate(uid = as.factor(uid),
               scale = as.factor(scale),
               dimension = substr(dimension, 1, 1) # makes it easier for EFA loadings display later
               ) %>%
        rename("response" = "response_reversed")
    return(data)
}

generic_gpt35_df <- process_data(file.path("data", "data_for_R", "generic_gpt35_df.csv"))
generic_gpt4_df <- process_data(file.path("data", "data_for_R", "generic_gpt4_df.csv"))

silicon_gpt35_df <- process_data(file.path("data", "data_for_R", "silicon_gpt35_df.csv"))
silicon_gpt4_df <- process_data(file.path("data", "data_for_R", "silicon_gpt4_df.csv"))
```

# Figure: First token of the responses

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_first_token_digit_item_frequencies.csv"),
                 show_col_types=F) %>%
  mutate(model = gsub("_relative_frequency", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model),
         first_token_is_digit = if_else(first_token_is_digit, "first token is digit", "first token is NOT digit")
         )
```

```{r fig.width=12}
p <- ggplot(data, aes(x = response_reversed, y = relative_frequency)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(relative_frequency)), 
                position = position_dodge(width = 0.9), 
                vjust = -0.3, 
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE)

p <- p + facet_grid(rows = vars(model), cols = vars(first_token_is_digit), scales = "free_x") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold", angle = 0)
  ) + 
  labs(y="Proportion", x="Item-level response") +
  scale_y_continuous(limits=c(0, 0.8))

p
# save_versioned_plot(p, "first_token_digit")
```

# Figure: internal consistency

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_cronbach_alpha_df.csv"), show_col_types=F) %>%
  pivot_longer(cols=-c("scale", "dimension"),
               names_to="model",
               values_to="alpha") %>%
  mutate(model = gsub("_alpha", "", model),
         scale = recode(scale, !!!scale_labels),
         model = recode(model, !!!model_labels),
         model = factor(model))
```

```{r fig.width=12, fig.height=9}
p <- data %>%
  ggplot(aes(x = dimension, y = alpha)) +
    geom_col(position = position_dodge()) +
    geom_hline(yintercept=0.7) +
    geom_text(aes(label = format_apa(alpha)), 
              position = position_dodge(width = 0.9), 
              vjust = -0.3, 
              color = "black", 
              size = 3.5, 
              check_overlap = TRUE) +
    scale_y_continuous(limits=c(0, 1.03)) + 
    facet_grid(rows = vars(model), cols = vars(scale), scales = "free_x") +
    theme(
      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.x = element_text(face = "bold", angle = 0),
      strip.text.y = element_text(face = "bold", angle = 0),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
    ) +
    labs(x="Dimension", y="Cronbach's alpha")

p
# save_versioned_plot(p, "cronbachs_alpha")
```







```{r}
calculate_alpha_for_group <- function(data) {
  tmp <- data %>%
    pivot_wider(id_cols=uid, 
                  names_from="item_index", names_prefix="item_", 
                  values_from="response") %>%
    select(-uid) %>%
    psych::alpha()
  
  return(tmp$total$std.alpha)
}

calculate_omega_for_group <- function(data) {
  tmp <- data %>%
    pivot_wider(id_cols=uid, 
                names_from="item_index", names_prefix="item_", 
                values_from="response") %>%
    select(-uid) %>%
    cor(use="pairwise") %>%
    as_tibble() %>%
    filter(!if_all(everything(), is.na)) %>%
    select(where(~!all(is.na(.)))) %>%
    omega(nfactors=1, fm="pa", rotate="oblimin")

  return(tmp$omega.tot)
}

calculate_glb_for_group <- function(data) {
  tmp <- data %>%
    pivot_wider(id_cols=uid, 
                names_from="item_index", names_prefix="item_", 
                values_from="response") %>%
    select(-uid) %>%
    cor(use="pairwise") %>%
    as_tibble() %>%
    filter(!if_all(everything(), is.na)) %>%
    select(where(~!all(is.na(.)))) %>%
    glb.fa()
  
  return(tmp$glb)
}

get_reliability_coeffs <- function(data, model_label) {
  data %>%
    mutate(dimension = case_when(
      scale=="BFI" & dimension=="O" ~ "Openness",
      scale=="BFI" & dimension=="C" ~ "Conscientiousness",
      scale=="BFI" & dimension=="E" ~ "Extraversion",
      scale=="BFI" & dimension=="A" ~ "Agreeableness",
      scale=="BFI" & dimension=="N" ~ "Neuroticism",
      scale=="BPAQ" & dimension=="A" ~ "Anger",
      scale=="BPAQ" & dimension=="H" ~ "Hostility",
      scale=="BPAQ" & dimension=="P" ~ "Physical",
      scale=="BPAQ" & dimension=="V" ~ "Verbal",
      scale=="PANAS" & dimension=="N" ~ "Negative",
      scale=="PANAS" & dimension=="P" ~ "Positive",
      scale=="SSCS" & dimension=="P" ~ "Personal Identity",
      scale=="SSCS" & dimension=="S" ~ "Self-efficacy",
    )) %>%
    group_by(scale, dimension) %>%
    summarise(model = model_label,
              alpha = calculate_alpha_for_group(cur_data()),
              omega = calculate_omega_for_group(cur_data()),
              glb = calculate_glb_for_group(cur_data()))
}

reliability_coeffs_df <- get_reliability_coeffs(generic_gpt35_df, "generic_gpt35") %>%
  bind_rows(get_reliability_coeffs(generic_gpt4_df, "generic_gpt4")) %>%
  bind_rows(get_reliability_coeffs(silicon_gpt35_df, "silicon_gpt35")) %>%
  bind_rows(get_reliability_coeffs(silicon_gpt4_df, "silicon_gpt4")) %>%
  pivot_longer(cols=-c(scale, dimension, model),
               names_to="coeff",
               values_to="value") %>%
  mutate(scale = recode(scale, !!!scale_labels),
         model = recode(model, !!!model_labels),
         model = factor(model))
reliability_coeffs_df
```

```{r fig.width=12, fig.height=9}
p <- reliability_coeffs_df %>%
  ggplot(aes(x = dimension, y = value, fill=coeff)) +
    geom_col(position = position_dodge(width=0.9)) +
    # geom_hline(yintercept=0.7) +
    geom_text(aes(label = format_apa(value)), 
              position = position_dodge(width = 0.9), 
              vjust = -0.3, 
              color = "black", 
              size = 2.5, 
              check_overlap = TRUE) +
    scale_y_continuous(limits=c(0, 1.03)) + 
    scale_fill_manual(values=c("black", "grey40", "grey80"), 
                      labels=c("Cronbach's alpha", 
                               "Greatest Lower Bound (glb)", 
                               "McDonald's omega")) +
    facet_grid(rows = vars(model), cols = vars(scale), scales = "free_x") +
    theme(
      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.x = element_text(face = "bold", angle = 0),
      strip.text.y = element_text(face = "bold", angle = 0),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
    ) +
    labs(x="Dimension", y="Reliability coefficient's value", fill="")

p
# save_versioned_plot(p, "reliability_coefficients")
```
# Figure: Criterion Validity

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_validity_df.csv"), show_col_types=F) %>%
  pivot_longer(cols=-c("var1", "var2"),
               names_to="model",
               values_to="r") %>%
  rowwise() %>%
  mutate(var1 = if_else(startsWith(var1, "BFI_"), strsplit(var1, "_")[[1]][2], var1)) %>%
  ungroup() %>%
  mutate(model = gsub("_r", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model),
         model = fct_relevel(model, levels(model)[3], after=0))
```

```{r fig.width=12, fig.height=9}
p <- data %>%
  ggplot(aes(x = var2, y = r)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(r),
                  vjust = if_else(r>0, -0.4, 1.2)), 
                position = position_dodge(width = 0.9),
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) + 
    scale_y_continuous(limits=c(-1.15,1.15)) +
    facet_grid(rows = vars(model), cols = vars(var1), scales = "free_x") +
    theme(
      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.x = element_text(face = "bold", size=10),
      strip.text.y = element_text(face = "bold", size=10, angle=0),
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size=10),
      panel.spacing.y = unit(0.3, "inches")
    ) + 
    labs(y="Correlation (Pearson's r)")

p
# save_versioned_plot(p, "validity")
```

# Figure: Intercorrelations

```{r}
data <- read.csv(file.path("data", "data_for_R", "intercorrs.csv")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model))
```

```{r fig.width=12, fig.height=9, warning=FALSE}
p_list <- list()
for (m in levels(data$model)) {
  p <- data %>%
    filter(model==m) %>%
    ggplot(aes(x = var1, y = var2, fill = r)) + 
      geom_tile() +
      geom_text(aes(label = if_else(is.na(r), NA, format_apa(r))), color = "black", size = 3) +
      scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                           midpoint = 0, limit = c(-1,1), space = "Lab", 
                           name="Correlation") +
      theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
            legend.position="top",
            legend.justification="center",
            legend.direction="vertical",
            legend.text=element_text(size=12)) +
      facet_wrap(~model, scales = "free") +
      labs(x = "", y = "") 
  p_list[[m]] <- p
}

legend <- get_legend(p_list[[1]])

p_list <- lapply(p_list, function(p) {
  p + theme(legend.position="none")
})

empty_plot <- ggplot() + theme_void() + theme(plot.background = element_blank(), panel.grid = element_blank(), panel.background = element_blank())

p <- ggarrange(
  ggarrange(empty_plot, empty_plot, p_list[[1]], empty_plot, legend, widths=c(0.1625, 0.1625, 0.35, 0.1625, 0.1625),
            nrow=1),
  ggarrange(empty_plot, p_list[[2]], empty_plot, p_list[[3]], empty_plot, widths=c(0.1, 0.35, 0.1, 0.35, 0.1), 
            nrow=1),
  ggarrange(empty_plot, p_list[[4]], empty_plot, p_list[[5]], empty_plot, widths=c(0.1, 0.35, 0.1, 0.35, 0.1), 
            nrow=1),
  nrow=3
)

p
# save_versioned_plot(p, "intercorrelations", bg="white")
```

# Figure: Item-level frequencies

```{r}
data <- read_csv(file.path("data", "data_for_R", "combined_bfi_item_frequencies.csv"),
                 show_col_types=F) %>%
  mutate(model = gsub("_relative_frequency", "", model),
         model = recode(model, !!!model_labels),
         model = factor(model))
```

```{r fig.width=12, fig.height=9}
p <- ggplot(data, aes(x = response_reversed, y = relative_frequency)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(label = format_apa(relative_frequency)), 
                position = position_dodge(width = 0.9),
                vjust = -0.3,
                color = "black", 
                size = 3.5, 
                check_overlap = TRUE) + 
    facet_grid(rows = vars(model), cols = vars(dimension), scales = "free_x") +
    theme(
      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.x = element_text(face = "bold"),
      strip.text.y = element_text(face = "bold", angle = 0)
    ) + 
    labs(y="Proportion", x="Item-level response") +
    scale_y_continuous(limits=c(0, 0.8))
p
# save_versioned_plot(p, "frequencies")
```

# EFA

```{r}
get_fa_model_loadings <- function(data, scale, model_label) {
  if (scale=="BFI") {
    data <- data %>%
      filter(scale=="BFI")
    n_factors <- 5
  } else if (scale=="SSCS") {
    data <- data %>%
      filter(scale=="SSCS")
    n_factors <- 2
  } else if (scale=="BPAQ") {
    data <- data %>%
      filter(scale=="BPAQ")
    n_factors <- 4
  } else if (scale=="PANAS") {
    data <- data %>%
      filter(scale=="PANAS")
    n_factors <- 2
  }
  
  items_no_variance <- c()
  if (scale=="BPAQ" & model_label=="silicon_gpt35") {
    items_no_variance <- c("A18", "V10")
  } else if (scale=="BPAQ" & model_label=="silicon_gpt4") {
    items_no_variance <- c("P8")
  } else if (scale=="SSCS" & model_label=="silicon_gpt35") {
    items_no_variance <- c("P10")
  }
  
  data <- data %>%
    select(-scale) %>%
    arrange(dimension, item_index) %>%
    pivot_wider(id_cols = uid,
                names_from = c("dimension", "item_index"),
                names_sep = "",
                values_from = "response") %>%
    select(-uid) %>% #for the cor_matrix
    select(-all_of(items_no_variance))
  
  cor_matrix <- cor(data, use = "pairwise.complete.obs")

  model <- fa(r = cor_matrix, nfactors = n_factors, 
                       rotate = "oblimin", fm = "pa", 
                       max.iter = 5000, n.obs = length(rownames(data)))
  
  df <- as.data.frame(unclass(model$loadings))
  df$item <- rownames(df)
  rownames(df) <- NULL
  df <- df %>%
    pivot_longer(cols=-item,
                 names_to="factor",
                 values_to="loading") %>%
    mutate(model=model_label)
  
  return(df)
}
```

## BFI

```{r}
if (!dir.exists(file.path("data", "bbc_data"))) {
  load(file.path("data", "bbc_data_for_sharing", "bbc_loadings.Rdata"))
} else {
  bbc_df <- process_data(file.path("data", "bbc_data", "bbc_df.csv"))
  bbc_loadings <- get_fa_model_loadings(bbc_df, "BFI", "bbc")
  save(bbc_loadings, file=file.path("data", "bbc_data_for_sharing", "bbc_loadings.Rdata"))
}
```

```{r}
efa_loadings_bfi <- bind_rows(bbc_loadings,
                          get_fa_model_loadings(generic_gpt35_df, "BFI", "generic_gpt35"),
                          get_fa_model_loadings(generic_gpt4_df, "BFI", "generic_gpt4"),
                          get_fa_model_loadings(silicon_gpt35_df, "BFI", "silicon_gpt35"),
                          get_fa_model_loadings(silicon_gpt4_df, "BFI", "silicon_gpt4")
                          ) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model))
```

### Figure EFA loadings

```{r fig.width=12, fig.height=6}
p <- efa_loadings_bfi %>%
  mutate(loading = abs(loading),
         item = factor(item, levels=rev(unique(efa_loadings_bfi$item)))) %>%
  ggplot(aes(x=factor, y=item, fill=loading)) +
      geom_tile() +
      scale_fill_gradient2(low="white", mid="darkgrey", high="black", midpoint=0.5, limit=c(0,1), space="Lab", name="Factor\nloading") +
      facet_wrap(~model, nrow=1) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(x = "Principal component", y = "Item index")

p
# save_versioned_plot(p, "efa_loadings_bfi", width=12, height=6, bg="white")
```

### Table EFA loadings

```{r}
efa_loadings_bfi %>%
  mutate(dimension = str_sub(item, 1, 1),
         item_index = as.double(str_sub(item, 2, -1))) %>%
  arrange(dimension, item_index, model, factor) %>%
  mutate(item=paste0(dimension, item_index)) %>%
  pivot_wider(id_cols=item,
              names_from=c("model", "factor"),
              values_from="loading") %>%
  mutate(across(is.numeric, ~format_apa(.))) #%>%
  # write.csv("figures-tables/efa_loadings.csv", row.names=F)
```

## All other questionnaires

### BPAQ

```{r}
efa_loadings_bpaq <- bind_rows(get_fa_model_loadings(generic_gpt35_df, "BPAQ", "generic_gpt35"),
                                get_fa_model_loadings(generic_gpt4_df, "BPAQ", "generic_gpt4"),
                                get_fa_model_loadings(silicon_gpt35_df, "BPAQ", "silicon_gpt35"),
                                get_fa_model_loadings(silicon_gpt4_df, "BPAQ", "silicon_gpt4")
                               ) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model))
```

#### Figure EFA loadings

```{r fig.width=12, fig.height=6}
p <- efa_loadings_bpaq %>%
  mutate(loading = abs(loading)) %>%
  # reordering is based on the item_index essentially
  ggplot(aes(x=factor, y=reorder(item, as.integer(str_sub(item, 2, -1)), decreasing=T), fill=loading)) +
      geom_tile() +
      scale_fill_gradient2(low="white", mid="darkgrey", high="black", midpoint=0.5, limit=c(0,1), space="Lab", name="Factor\nloading") +
      facet_wrap(~model, nrow=1) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(x = "Principal component", y = "Item index")

p
# save_versioned_plot(p, "efa_loadings_bpaq", width=12, height=6, bg="white")
```

### SSCS

```{r}
efa_loadings_sscs <- bind_rows(get_fa_model_loadings(generic_gpt35_df, "SSCS", "generic_gpt35"),
                                get_fa_model_loadings(generic_gpt4_df, "SSCS", "generic_gpt4"),
                                get_fa_model_loadings(silicon_gpt35_df, "SSCS", "silicon_gpt35"),
                                get_fa_model_loadings(silicon_gpt4_df, "SSCS", "silicon_gpt4")
                               ) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model))
```

#### Figure EFA loadings

```{r fig.width=12, fig.height=6}
p <- efa_loadings_sscs %>%
  mutate(loading = abs(loading),
         item = factor(item, levels=rev(unique(efa_loadings_sscs$item)))) %>%
  ggplot(aes(x=factor, y=item, fill=loading)) +
      geom_tile() +
      scale_fill_gradient2(low="white", mid="darkgrey", high="black", midpoint=0.5, limit=c(0,1), space="Lab", name="Factor\nloading") +
      facet_wrap(~model, nrow=1) +
      theme_minimal() +
      # theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(x = "Principal component", y = "Item index")

p
# save_versioned_plot(p, "efa_loadings_sscs", width=12, height=6, bg="white")
```

### PANAS

```{r}
efa_loadings_panas <- bind_rows(get_fa_model_loadings(generic_gpt35_df, "PANAS", "generic_gpt35"),
                                get_fa_model_loadings(generic_gpt4_df, "PANAS", "generic_gpt4"),
                                get_fa_model_loadings(silicon_gpt35_df, "PANAS", "silicon_gpt35"),
                                get_fa_model_loadings(silicon_gpt4_df, "PANAS", "silicon_gpt4")
                               ) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model))
```

#### Figure EFA loadings

```{r fig.width=12, fig.height=6}
p <- efa_loadings_panas %>%
  mutate(loading = abs(loading),
         item = factor(item, levels=rev(unique(efa_loadings_panas$item)))) %>%
  ggplot(aes(x=factor, y=item, fill=loading)) +
      geom_tile() +
      scale_fill_gradient2(low="white", mid="darkgrey", high="black", midpoint=0.5, limit=c(0,1), space="Lab", name="Factor\nloading") +
      facet_wrap(~model, nrow=1) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(x = "Principal component", y = "Item index")

p
# save_versioned_plot(p, "efa_loadings_panas", width=12, height=6, bg="white")
```

# CFA

```{r}
# Function to remove specific variables from a lavaan-like cfa_spec
# Courtesy of ChatGPT
remove_vars_from_cfa <- function(cfa_spec, vars_to_remove) {
  lines <- unlist(strsplit(cfa_spec, "\n"))
  updated_lines <- sapply(lines, function(line) {
    for (var in vars_to_remove) {
      line <- gsub(paste0(" \\+ ", var), "", line)
      line <- gsub(paste0(var, " \\+ "), "", line)
    }
    return(line)
  })
  updated_cfa_spec <- paste(updated_lines, collapse = "\n")
  return(updated_cfa_spec)
}

get_cfa_fit <- function(scale_label, cfa_model, data, items_no_variance=c()) {
  if (length(items_no_variance)>0) {
    cfa_model <- remove_vars_from_cfa(cfa_model, items_no_variance)
  }
  
  data <- data %>%
    filter(scale==scale_label) %>%
    pivot_wider(id_cols="uid",
                names_from=c("dimension", "item_index"),
                names_sep="",
                values_from="response") %>%
    select(-all_of(items_no_variance))
  
  fit <- cfa(cfa_model, data=data, std.lv=T, estimator="MLM")
  
  return(fit)
}

get_loadings_from_cfa_fit <- function(fit, model_label) {
  parameterEstimates(fit) %>%
    filter(op=="=~") %>%
    select(lhs, rhs, est) %>%
    rename('latent'=lhs, 'item'=rhs) %>%
    mutate(model=model_label) %>%
    as_tibble()
}

get_covariances_from_model_fit <- function(fit, model_label) {
  latent_variables <- parameterEstimates(fit) %>%
    filter(lhs==rhs, est==1) %>%
    pull(lhs)
  
  parameterEstimates(fit) %>%
    filter(op=="~~", est < 1, lhs %in% latent_variables) %>%
    select(lhs, rhs, est) %>%
    mutate(model=model_label) %>%
    as_tibble()
}

get_fit_indices <- function(fit, model_label) {
  measures <- fitMeasures(fit, fit.measures = c("GFI", "IFI", "RMSEA"))
  tibble(model = model_label,
         gfi = measures["gfi"],
         ifi = measures["ifi"],
         rmsea = measures["rmsea"])
}
```

## BFI

```{r}
cfa_model_bfi <- '
  O =~ O5 + O10 + O15 + O20 + O25 + O30 + O35 + O40 + O41 + O44
  C =~ C3 + C8 + C13 + C18 + C23 + C28 + C33 + C38 + C43
  E =~ E1 + E6 +E11 + E16 + E21 + E26 + E31 + E36
  A =~ A2 + A7 + A12 + A17 + A22 + A27 + A32 + A37 + A42
  N =~ N4 + N9 + N14 + N19 + N24 + N29 + N34 + N39
'
```

```{r}
if (!dir.exists(file.path("data", "bbc_data"))) {
  load(file.path("data", "bbc_data_for_sharing", "cfa_fit_bfi_bbc.Rdata"))
} else {
  bbc_df <- process_data(file.path("data", "bbc_data", "bbc_df.csv"))
  cfa_fit_bfi_bbc <- get_cfa_fit("BFI", cfa_model_bfi, bbc_df)
  save(cfa_fit_bfi_bbc, file=file.path("data", "bbc_data_for_sharing", "cfa_fit_bfi_bbc.Rdata"))
}

cfa_fit_bfi_generic_gpt35 <- get_cfa_fit("BFI", cfa_model_bfi, generic_gpt35_df)
cfa_fit_bfi_generic_gpt4 <- get_cfa_fit("BFI", cfa_model_bfi, generic_gpt4_df)
cfa_fit_bfi_silicon_gpt35 <- get_cfa_fit("BFI", cfa_model_bfi, silicon_gpt35_df)
# below one - bfi silicon gpt4 - is not positive definite so is not ran
# cfa_fit_bfi_silicon_gpt4 <- get_cfa_fit("BFI", cfa_model_bfi, silicon_gpt4_df)
```

```{r}
data <- get_loadings_from_cfa_fit(cfa_fit_bfi_bbc, "bbc") %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_bfi_generic_gpt35, "generic_gpt35")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_bfi_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_bfi_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         item = factor(item, levels=rev(unique(cur_data()$item))),
         latent = case_when(
           latent=="A" ~ "Agreeableness",
           latent=="O" ~ "Openness",
           latent=="N" ~ "Neuroticism",
           latent=="C" ~ "Conscientiousness",
           latent=="E" ~ "Extraversion"
         ),
         latent = factor(latent, 
                         levels=c("Openness", "Conscientiousness",
                                  "Extraversion", "Agreeableness", "Neuroticism")))
```

```{r fig.width=12, fig.height=9}
p1 <- data %>%
  ggplot(aes(x=latent, y=item, fill=abs(est))) +
    geom_tile() +
    geom_text(aes(label=format_apa(est),
                  color=if_else(abs(est)>.75, "white", "black")),
              size=3,
              position=position_identity()) +
    scale_color_identity() + # necessary for the geom_text color to work
    scale_fill_gradient2(low="white", mid="darkgrey", high="black",
                         midpoint=0.5, limit=c(0,1), space="Lab",
                         name="Standardized\nloading") +
    facet_wrap(~model, nrow=1) +
    theme_minimal() +
    labs(x = "Latent factor", y = "Item") +
    theme(axis.text.x = element_text(angle=45, hjust=1))

p1
```

```{r}
data <- get_covariances_from_model_fit(cfa_fit_bfi_bbc, "bbc") %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_bfi_generic_gpt35, "generic_gpt35")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_bfi_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_bfi_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         lhs = case_when(
           lhs=="A" ~ "Agreeableness",
           lhs=="O" ~ "Openness",
           lhs=="N" ~ "Neuroticism",
           lhs=="C" ~ "Conscientiousness",
           lhs=="E" ~ "Extraversion"
         ),
         lhs = factor(lhs, 
                         levels=c("Openness", "Conscientiousness",
                                  "Extraversion", "Agreeableness", "Neuroticism")),
         rhs = case_when(
           rhs=="A" ~ "Agreeableness",
           rhs=="O" ~ "Openness",
           rhs=="N" ~ "Neuroticism",
           rhs=="C" ~ "Conscientiousness",
           rhs=="E" ~ "Extraversion"
         ),
         rhs = factor(rhs, 
                         levels=c("Openness", "Conscientiousness",
                                  "Extraversion", "Agreeableness", "Neuroticism")))
```
      
```{r fig.width=12, fig.height=3}
p2 <- data %>%
  ggplot(aes(x=lhs, y=rhs, fill=est)) +
  geom_tile() +
  geom_text(aes(label = if_else(is.na(est), NA, format_apa(est))), color = "black", size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                           midpoint = 0, limit = c(-1,1), space = "Lab", 
                           name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
            # legend.position="top",
            # legend.justification="center",
            # legend.direction="vertical",
            # legend.text=element_text(size=12)
        ) +
  facet_wrap(~model, nrow=1) +
  labs(x = "", y = "")

p2
```

```{r fig.width=12, fig.height=12}
p <- ggarrange(p1, p2,
          labels=c("A", "B"),
          heights=c(9,3),
          ncol=1)

p
# save_versioned_plot(p, "cfa_results_bfi", width=12, height=12, bg="white")
```

```{r}
get_fit_indices(cfa_fit_bfi_bbc, "bbc") %>%
  bind_rows(get_fit_indices(cfa_fit_bfi_generic_gpt35, "generic_gpt35")) %>%
  bind_rows(get_fit_indices(cfa_fit_bfi_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_fit_indices(cfa_fit_bfi_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         across(is.numeric, ~format_apa(.)))
```

## All other questionnaires

### BPAQ

```{r}
cfa_model_bpaq <- '
  P =~ P1 + P2 + P3 + P4 + P5 + P6 + P7 + P8 + P9
  V =~ V10 + V11 + V12 + V13 + V14
  A =~ A15 + A16 + A17 + A18 + A19 + A20 + A21
  H =~ H22 + H23 + H24 + H25 + H26 + H27 + H28 + H29
'
```

```{r}
cfa_fit_bpaq_generic_gpt35 <- get_cfa_fit("BPAQ", cfa_model_bpaq, generic_gpt35_df)
cfa_fit_bpaq_generic_gpt4 <- get_cfa_fit("BPAQ", cfa_model_bpaq, generic_gpt4_df)
cfa_fit_bpaq_silicon_gpt35 <- get_cfa_fit("BPAQ", cfa_model_bpaq, silicon_gpt35_df, c("A18", "V10"))
# below one - bpaq silicon gpt4 - is not positive definite so is not ran
# cfa_fit_bpaq_silicon_gpt4 <- get_cfa_fit("BPAQ", cfa_model_bpaq, silicon_gpt4_df, c("P8"))
```

```{r}
data <- get_loadings_from_cfa_fit(cfa_fit_bpaq_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_bpaq_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_bpaq_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         item = factor(item, levels=rev(unique(cur_data()$item))),
         latent = case_when(
           latent=="A" ~ "Anger",
           latent=="P" ~ "Physical",
           latent=="V" ~ "Verbal",
           latent=="H" ~ "Hostility",
         ),
         latent = factor(latent, levels=c("Physical", "Verbal", "Anger", "Hostility")))
```

```{r fig.width=12, fig.height=9}
p1 <- data %>%
  ggplot(aes(x=latent, y=item, fill=abs(est))) +
    geom_tile() +
    geom_text(aes(label=format_apa(est),
                  color=if_else(abs(est)>.75, "white", "black")),
              position=position_identity()) +
    scale_fill_gradient2(low="white", mid="darkgrey", high="black",
                         midpoint=0.5, limit=c(0,1), space="Lab",
                         name="Standardized\nloading") +
    scale_color_identity() + # necessary for the geom_text color to work
    facet_wrap(~model, nrow=1) +
    theme_minimal() +
    labs(x = "Latent factor", y = "Item")

p1
```

```{r}
data <- get_covariances_from_model_fit(cfa_fit_bpaq_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_bpaq_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_bpaq_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         lhs = case_when(
           lhs=="A" ~ "Anger",
           lhs=="P" ~ "Physical",
           lhs=="V" ~ "Verbal",
           lhs=="H" ~ "Hostility",
         ),
         lhs = factor(lhs, levels=c("Physical", "Verbal", "Anger", "Hostility")),
         rhs = case_when(
           rhs=="A" ~ "Anger",
           rhs=="P" ~ "Physical",
           rhs=="V" ~ "Verbal",
           rhs=="H" ~ "Hostility",
         ),
         rhs = factor(rhs, levels=c("Physical", "Verbal", "Anger", "Hostility")))
```
      
```{r fig.width=12, fig.height=3}
p2 <- data %>%
  ggplot(aes(x=lhs, y=rhs, fill=est)) +
  geom_tile() +
  geom_text(aes(label = if_else(is.na(est), NA, format_apa(est))), color = "black", size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                           midpoint = 0, limit = c(-1,1), space = "Lab", 
                           name="Correlation") +
  theme_minimal() +
  facet_wrap(~model, nrow=1) +
  labs(x = "", y = "")

p2
```

```{r fig.width=12, fig.height=12}
p <- ggarrange(p1, p2,
          labels=c("A", "B"),
          heights=c(9,3),
          ncol=1)

p
# save_versioned_plot(p, "cfa_results_bpaq", width=12, height=12, bg="white")
```
```{r}
get_fit_indices(cfa_fit_bpaq_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_fit_indices(cfa_fit_bpaq_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_fit_indices(cfa_fit_bpaq_silicon_gpt35, "silicon_gpt35")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         across(is.numeric, ~format_apa(.)))
```

### SSCS

```{r}
cfa_model_sscs <- '
  P =~ P1 + P2 + P7 + P10 + P11
  S =~ S3 + S4 + S5 + S6 +S8 + S9
'
```

```{r}
cfa_fit_sscs_generic_gpt35 <- get_cfa_fit("SSCS", cfa_model_sscs, generic_gpt35_df)
cfa_fit_sscs_generic_gpt4 <- get_cfa_fit("SSCS", cfa_model_sscs, generic_gpt4_df)
cfa_fit_sscs_silicon_gpt35 <- get_cfa_fit("SSCS", cfa_model_sscs, silicon_gpt35_df, c("P10"))
cfa_fit_sscs_silicon_gpt4 <- get_cfa_fit("SSCS", cfa_model_sscs, silicon_gpt4_df)
```

```{r}
data <- get_loadings_from_cfa_fit(cfa_fit_sscs_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_sscs_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_sscs_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_sscs_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         item = factor(item, levels=rev(unique(cur_data()$item))),
         latent = case_when(
           latent=="P" ~ "Personal Identity",
           latent=="S" ~ "Self-efficacy",
         ),
         latent = factor(latent, levels=c("Personal Identity", "Self-efficacy")))
```

```{r fig.width=12, fig.height=9}
p1 <- data %>%
  ggplot(aes(x=latent, y=item, fill=abs(est))) +
    geom_tile() +
    geom_text(aes(label=format_apa(est),
                  color=if_else(abs(est)>.75, "white", "black")),
              position=position_identity()) +
    scale_fill_gradient2(low="white", mid="darkgrey", high="black",
                         midpoint=0.5, limit=c(0,1), space="Lab",
                         name="Standardized\nloading") +
    scale_color_identity() + # necessary for the geom_text color to work
    facet_wrap(~model, nrow=1) +
    theme_minimal() +
    labs(x = "Latent factor", y = "Item")

p1
```

```{r}
data <- get_covariances_from_model_fit(cfa_fit_sscs_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_sscs_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_sscs_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_sscs_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         lhs = case_when(
           lhs=="P" ~ "Personal Identity",
           lhs=="S" ~ "Self-efficacy",
         ),
         lhs = factor(lhs, levels=c("Personal Identity", "Self-efficacy")),
         rhs = case_when(
           rhs=="P" ~ "Personal Identity",
           rhs=="S" ~ "Self-efficacy",
         ),
         rhs = factor(rhs, levels=c("Personal Identity", "Self-efficacy")))
```
      
```{r fig.width=12, fig.height=3}
p2 <- data %>%
  ggplot(aes(x=lhs, y=rhs, fill=est)) +
  geom_tile() +
  geom_text(aes(label = if_else(is.na(est), NA, format_apa(est))), color = "black", size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                           midpoint = 0, limit = c(-1,1), space = "Lab", 
                           name="Correlation") +
  theme_minimal() +
  facet_wrap(~model, nrow=1) +
  labs(x = "", y = "")

p2
```

```{r fig.width=12, fig.height=12}
p <- ggarrange(p1, p2,
          labels=c("A", "B"),
          heights=c(9,3),
          ncol=1)

p
# save_versioned_plot(p, "cfa_results_sscs", width=12, height=12, bg="white")
```
```{r}
get_fit_indices(cfa_fit_sscs_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_fit_indices(cfa_fit_sscs_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_fit_indices(cfa_fit_sscs_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_fit_indices(cfa_fit_sscs_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         across(is.numeric, ~format_apa(.)))
```

### PANAS

```{r}
cfa_model_panas <- '
  P =~ P1 + P3 + P5 + P9 + P10 + P12 + P14 + P16 + P17 + P19
  N =~ N2 + N4 + N6 + N7 + N8 + N11 + N13 + N15 + N18 + N20
'
```

```{r}
cfa_fit_panas_generic_gpt35 <- get_cfa_fit("PANAS", cfa_model_panas, generic_gpt35_df)
cfa_fit_panas_generic_gpt4 <- get_cfa_fit("PANAS", cfa_model_panas, generic_gpt4_df)
cfa_fit_panas_silicon_gpt35 <- get_cfa_fit("PANAS", cfa_model_panas, silicon_gpt35_df)
cfa_fit_panas_silicon_gpt4 <- get_cfa_fit("PANAS", cfa_model_panas, silicon_gpt4_df)
```

```{r}
data <- get_loadings_from_cfa_fit(cfa_fit_panas_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_panas_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_panas_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_loadings_from_cfa_fit(cfa_fit_panas_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         item = factor(item, levels=rev(unique(cur_data()$item))),
         latent = case_when(
           latent=="P" ~ "Positive",
           latent=="N" ~ "Negative",
         ),
         latent = factor(latent, levels=c("Positive", "Negative")))
```

```{r fig.width=12, fig.height=9}
p1 <- data %>%
  ggplot(aes(x=latent, y=item, fill=abs(est))) +
    geom_tile() +
    geom_text(aes(label=format_apa(est),
                  color=if_else(abs(est)>.75, "white", "black")),
              position=position_identity()) +
    scale_fill_gradient2(low="white", mid="darkgrey", high="black",
                         midpoint=0.5, limit=c(0,1), space="Lab",
                         name="Standardized\nloading") +
    scale_color_identity() + # necessary for the geom_text color to work
    facet_wrap(~model, nrow=1) +
    theme_minimal() +
    labs(x = "Latent factor", y = "Item")
```

```{r}
data <- get_covariances_from_model_fit(cfa_fit_panas_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_panas_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_panas_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_covariances_from_model_fit(cfa_fit_panas_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         lhs = case_when(
           lhs=="P" ~ "Positive",
           lhs=="N" ~ "Negative",
         ),
         lhs = factor(lhs, levels=c("Positive", "Negative")),
         rhs = case_when(
           rhs=="P" ~ "Positive",
           rhs=="N" ~ "Negative",
         ),
         rhs = factor(rhs, levels=c("Positive", "Negative")))
```
      
```{r fig.width=12, fig.height=3}
p2 <- data %>%
  ggplot(aes(x=lhs, y=rhs, fill=est)) +
  geom_tile() +
  geom_text(aes(label = if_else(is.na(est), NA, format_apa(est))), color = "black", size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                           midpoint = 0, limit = c(-1,1), space = "Lab", 
                           name="Correlation") +
  theme_minimal() +
  facet_wrap(~model, nrow=1) +
  labs(x = "", y = "")
```

```{r fig.width=12, fig.height=12}
p <- ggarrange(p1, p2,
          labels=c("A", "B"),
          heights=c(9,3),
          ncol=1)

p
# save_versioned_plot(p, "cfa_results_panas", width=12, height=12, bg="white")
```

```{r}
get_fit_indices(cfa_fit_panas_generic_gpt35, "generic_gpt35") %>%
  bind_rows(get_fit_indices(cfa_fit_panas_generic_gpt4, "generic_gpt4")) %>%
  bind_rows(get_fit_indices(cfa_fit_panas_silicon_gpt35, "silicon_gpt35")) %>%
  bind_rows(get_fit_indices(cfa_fit_panas_silicon_gpt4, "silicon_gpt4")) %>%
  mutate(model = recode(model, !!!model_labels),
         model = factor(model),
         across(is.numeric, ~format_apa(.)))
```

# Figure: Trait bias

```{r}
data <- read_csv(file.path("data", "data_for_R", "bias_df.csv"), show_col_types=F) %>%
  mutate(model = recode(model, !!!model_labels2))
```

```{r fig.width=12, fig.height=6, warning=FALSE}
data_summary_stats_labels <- data %>%
  group_by(model, trait) %>%
  summarise(label = paste0('<i>M</i>=', format_apa(mean(bias)), ' (<i>SD</i>=', format_apa(sd(bias)), ')'))

data_p_val_labels <- data %>%
  group_by(trait) %>%
  summarise(label = if_else(
    t.test(bias~model, data=cur_data())$p.value < 0.05,
    paste0('<b><i>p</i> ', format_apa(t.test(bias~model, data=cur_data())$p.value, T), '</b>'),
    paste0('<i>p</i> ', format_apa(t.test(bias~model, data=cur_data())$p.value, T))
    )
  )


p <- ggplot(data, aes(x=model, y=bias)) +
    geom_violin(width=0.6) +
    geom_boxplot(width=0.05, alpha=0.2) +
    ggtext::geom_richtext(data=data_summary_stats_labels, 
                            mapping=aes(y=3.5, x=model, label=label, group=NULL, fill=NULL), 
                            size=3,
                            show.legend=F) +
    ggtext::geom_richtext(data=data_p_val_labels, 
                            mapping=aes(y=3, x=1.5, label=label, group=NULL, fill=NULL), 
                            size=3,
                            show.legend=F) +
    labs(x="Model", y="Trait bias") +
    facet_wrap(~trait) +
    scale_y_continuous(limits=c(0, 3.7))

p
# save_versioned_plot(p, "trait_bias", width=12, height=6)
```

# Figure: Bias correlations

```{r}
data <- read_csv(file.path("data", "data_for_R", "bias_corrs_df.csv"), show_col_types=F)
```

```{r}
labels_mapper_p1 <- c(
  "BFI_Agreeableness" = "Agreeableness", "BFI_Conscientiousness" = "Conscientiousness",
  "BFI_Extraversion" = "Extraversion", "BFI_Openness" = "Openness", "BFI_Neuroticism" = "Neuroticism"
)
labels_mapper_p2 <- c(
  "age" = "Age", "sex" = "Sex", "ethnic" = "Ethnicity", "income" = "Income",
  "rstat_1" = "Relationship status", "n_sib" = "Number of siblings"
)
labels_mapper_p3 <- c(
  "st_pub" = "State or private schooling", "chldrn" = "Number of children", 
  "occ_sta" = "Occupational status", "occ_cat" = "Occupation category",
  "m_schl" = "Mother's education", "f_schl" = "Father's education"
)
```

```{r fig.width=12, fig.height=9}
p1 <- data %>%
  filter(str_detect(var, "^BFI_")) %>%
  mutate(var = recode(var, !!!labels_mapper_p1)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1.1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="Correlation coefficient", fill="") +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r fig.width=12, fig.height=9}
p2_vars <- c("age", "sex", "ethnic", "income", "rstat_1", "n_sib")

labels_mapper <- c("st_pub" = "State or private schooling", "chldrn" = "Number of children", 
                   "occ_sta" = "Occupational status", "occ_cat" = "Occupation category",
                   "m_schl" = "Mother's education", "f_schl" = "Father's education")

p2 <- data %>%
  filter(var %in% p2_vars) %>%
  mutate(var = recode(var, !!!labels_mapper_p2)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1.1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="", fill="") +
      guides(fill="none") +
      scale_y_continuous(limits=c(-0.05, 0.11)) +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3 <- data %>%
  filter(!str_detect(var, "^BFI_"), !var %in% p2_vars) %>%
  mutate(var = recode(var, !!!labels_mapper_p3)) %>%
  ggplot(aes(x = var, y = r, fill=model)) +
      geom_col(position = position_dodge()) +
      geom_text(aes(label = format_apa(r),
                    vjust = if_else(r>0, -0.3, 1.1)), 
                  position = position_dodge(width = 0.9), 
                  color = "black", 
                  size = 3.5, 
                  check_overlap = TRUE) +
      labs(x="", y="", fill="") +
      scale_fill_grey(labels=c("GPT-3.5", "GPT-4")) +
      scale_y_continuous(limits=c(-0.05, 0.11)) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r fig.width=12, fig.height=9}
p2_p3 <- ggarrange(p2, p3, nrow=2, ncol=1, common.legend = T, legend="none")
```

```{r fig.width=12, fig.height=9}
p <- ggarrange(p1, p2_p3,
          font.label=list(size=20, color="black", face="bold"),
          nrow=1, ncol=2,
          common.legend=T,
          legend="bottom")

p
# save_versioned_plot(p, "trait_bias_correlations", bg="white")
```

```{r}
knitr::knit_exit()
```

# Exit



























